Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-21 forecast/paper/supplement-text.tex
minor changes

The goal of this section is to quantify the effect of
not properly accounting for the question of "what was known when" in
performing retrospective evaluations of forecasters.   Figures
\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show 
what Figures 3 and 4 in the main paper would have looked like if we
had simply trained all models using the finalized data rather than
using vintage data.  This comparison can be seen more
straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
\ref{fig:hot-honest-v-finalized}, which show the ratio in performance between
the vintage and finalized 
versions.  When methods are 
given the finalized version of the data rather than the version available at the
time that the forecast would have been made, all methods 
appear (misleadingly) to have better performance than they would
have had if run prospectively.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
achieved using finalized data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that longer-horizon forecasters rely less heavily on recent data than very
short-horizon forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

@@

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
rates $Y_{\ell,t}$ is defined as
\[
\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
\]
where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
are the sample mean and
standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
sequence of Pearson correlations for each selected $a$, where we align
$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
days in the future. In this case we say that $X_{\ell,t}$ is lagging
$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
$X_{\ell,t}$ leads $Y_{\ell,t}$.

@@

Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
manuscript) there is no obvious relationship. This is heartening as laggingness
should not aid forecasting performance. On the other hand, if an indicator is
more lagging than it is leading, this may suggest diminished performance.
Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
leadingness and laggingness with the difference in WIS. The pattern here is
largely similar to the pattern in leadingness described in the manuscript: the
relationship is strongest in down periods and weakest in up periods with the
strength diminishing as we move from down to flat to up for all indicators.

>>>>

The goal of this section is to quantify the effect of
not properly accounting for the question of ``what was known when'' in
performing retrospective evaluations of forecasters.   Figures
\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show 
what Figures 3 and 4 in the main paper would have looked like if we
had simply trained all models using the finalized data rather than
using vintage data.  This comparison can be seen more
straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
\ref{fig:hot-honest-v-finalized}, which show the ratio in performance between
the vintage and finalized 
versions.  When methods are 
given the finalized version of the data rather than the version available at the
time that the forecast would have been made, all methods 
appear (misleadingly) to have better performance than they would
have had if run prospectively.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
achieved using finalized data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that longer-horizon forecasters rely less heavily on recent data than very
short-horizon forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

@@

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The CCF of an indicator $X_{\ell,t}$ and case
rates $Y_{\ell,t}$ is defined as
\[
\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
\]
where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
are the sample mean and
standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
sequence of Pearson correlations for each selected $a$, where we align
$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
days in the future. In this case we say that $X_{\ell,t}$ is lagging
$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
$X_{\ell,t}$ leads $Y_{\ell,t}$.

@@

Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
manuscript) there is no obvious relationship that holds consistently
across indicators. This is heartening as laggingness
should not aid forecasting performance. On the other hand, if an indicator is
more lagging than it is leading, this may suggest diminished performance.
Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
leadingness and laggingness with the difference in WIS. The pattern here is
largely similar to the pattern in leadingness described in the manuscript: the
relationship is strongest in down periods and weakest in up periods with the
strength diminishing as we move from down to flat to up for all indicators.
